---
title: Análisis exploratorio de un conjunto de datos de indicadores urbanos
author:
  - name: Rebeca Company
    affil: 1,2,\dagger
  - name: Alejandro Dionis
    affil: 1,3,\dagger
  - name: Gabriel Ivars
    affil: 1,4,\dagger
  - name: Julio Garcia
    affil: 1,5,\dagger
affiliation:
  - num: 1
    address: |
      Universitat de València -
      ETSE-UV, Avinguda de l'Universitat, 46100 Burjassot, Valencia
  - num: 2
    address: |
      Máster Universitario en Ciencia de Datos
    email: recombar@alumni.uv.es
  - num: 3
    address: |
      Máster Universitario en Ciencia de Datos
    email: adioros@alumni.uv.es
  - num: 4
    address: |
      Máster Universitario en Ciencia de Datos
    email: gaia2@alumni.uv.es
  - num: 5
    address: |
      Máster Universitario en Ciencia de Datos
    email: jugarbus@alumni.uv.es    
# author citation list in chicago format
authorcitation: |
  Company, R.; Dionis, A.; Ivars, G.; Garcia, J.
# firstnote to eighthnote
firstnote: |
  Todos los autores contribuyeron de manera equitativa a este trabajo.
correspondence: |
  etse@uv.es; Tel.: +34-963-54-3211.
  
# document options
journal: Universitat de València
type: article
status: submit
# front matter
simplesummary: |
  El presente artículo es un proyecto de la Asignatura Análisis Exploratorio de Datos, en el que se realiza un análisis exploratorio del conjunto de datos de Indicadores Urbanos, publicado por el Instituto Nacional de Estadística.
abstract: |
  REDACTAR MÁS TARDE
# back matter
keywords: |
  Indicadores Urbanos; Tasa ; Población; REVISAR MAS TARDE
authorcontributions: |
    Todos los autores contribuyeron de manera equitativa a este trabajo.
funding: |
  This research received no external funding.
dataavailability: |
  Los conjuntos de datos analizados en el presente artículo están disponibles en la sección "Indicadores Urbanos. Últimos datos." del repositorio público de conjuntos de datos abiertos del Instituto Nacional de Estadística INEbase
  \url{https://www.ine.es/dyngs/INEbase/listaoperaciones.htm}.
conflictsofinterest: |
  Los autores declaran no tener ningún conflicto de interés
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: INE
    long: Instituto Nacional de Estadística
  - short: RMT
    long: REVISAR MAS TARDE
bibliography: mybibfile.bib
appendix: appendix.tex
# To use endnotes, change endnotes: true , then use \endnote{This 28 is an endnote.}.
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

# Introduction

The introduction should briefly place the study in a broad context and highlight
why it is important. It should define the purpose of the work and its
significance. The current state of the research field should be reviewed
carefully and key publications cited. Please highlight controversial and
diverging hypotheses when necessary. Finally, briefly mention the main aim of
the work and highlight the principal conclusions. As far as possible, please
keep the introduction comprehensible to scientists outside your particular
field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998;
@leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008.
Some MDPI journals use Chicago and others use APA, this template should choose
the correct citation format for you once you specify the journal in the YAML
header.




# Materials and Methods

Materials and Methods should be described with sufficient details to allow 
others to replicate and build on published results. Please note that publication 
of your manuscript implicates that you must make all materials, data, computer 
code, and protocols associated with the publication available to readers. Please
disclose at the submission stage any restrictions on the availability of 
materials or information. New methods and protocols should be described in 
detail while well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a publicly 
available database should specify where the data have been deposited and provide
the relevant accession numbers. If the accession numbers have not yet been
obtained at the time of submission, please state that they will be provided
during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require
ethical approval must list the authority that provided approval and the
corresponding ethical approval code.

# Results

This section may be divided by subheadings. It should provide a concise and
precise description of the experimental results, their interpretation as well
as the experimental conclusions that can be drawn.



# Conclusion

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.


# Como utilizar la plantilla

## Notas de pie de página

To use endnotes, change `endnotes: true` in the YAML header, then use
`\endnote{This is an endnote.}`.

## Código en el texto:

Una forma 

```yaml
title: Title of the paper
author:
  - name: first and last name
    affil: |
      One or more comma seperated numbers corresponding to affilitation
      and one or more  comma seperated symbols corresponding 
      optional notes.
    orcid: optional orcid number
affiliation:  
  - num: 1,..., n for each affiliation
    address: required
    email: required
authorcitation: |
  Lastname, F.
```
Otra forma de insertar código:

The `status` variable should generally not be changed by authors. The `type` variable describes the type of of submission and defaults to `article` but

Mostrar los chunks:

```{r mdpitype, echo = TRUE}
type <- c("abstract, addendum, article, book, bookreview, briefreport, 
casereport, comment, commentary, communication, conferenceproceedings, 
correction, conferencereport, entry, expressionofconcern, extendedabstract, 
datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, 
obituary, opinion, projectreport, reply, retraction, review, perspective, 
protocol, shortnote, studyprotocol, systematicreview, supfile, 
technicalnote, viewpoint, guidelines, registeredreport, tutorial")
knitr::kable(rticles::string_to_table(type, 3), 
             col.names = NULL, format = "latex", booktab = TRUE, 
             caption = "MDPI article types.")
```

## Tablas

Se citan con el orden en el que aparece el código, no el texto

Table \ref{tab:mdpinames}.  can be replaced with any of the ones in Table \ref{tab:mdpitype}


## Salto de página a una página con orientación horizontal

\startlandscape
```{r, mdpinames, echo=FALSE, results='asis'}
mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis"

knitr::kable(rticles::string_to_table(mdpi_journals, 8),
             col.names = NULL,
             format = "latex",
             booktabs = TRUE, 
             caption = "MDPI journal names.",
             longtable = TRUE)

```
\finishlandscape

## Listas

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

## Figures, Tables and Schemes

All figures and tables should be cited in the main text as Figure \ref{fig:fig1}, 
\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks 
include the `\\label{}` tag in the `fig.cap` attribute of the R chunk: 
`fig.cap = "Fancy Caption\\label{fig:plot}"`.

```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```


When making tables using `kable`, it is suggested to use
the `format="latex"` and `tabl.envir="table"` arguments
to ensure table numbering and compatibility with the mdpi
document class.

```{r tab1, echo=FALSE}
knitr::kable(mtcars[1:5, 1:3], format = "latex", 
             booktabs = TRUE, 
             caption = "This is a table caption. Tables should be placed in the 
             main text near to the first time they are~cited.", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```



For a very wide table, landscape layouts are allowed.


\startlandscape

```{r tab2, echo=FALSE}
df <- data.frame(`Title 1` = c("Entry 1", "Entry 2"),
                 `Title 2` = c("Data", "Data"),
                 `Title 3` = c("Data", "Data"),
                 `Title 4` = c("This cell has some longer content that runs over
                               two lines",
                               "Data"))
knitr::kable(df, format = "latex", 
             booktabs = TRUE, 
             caption = "This is a very wide table", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```

\finishlandscape

## Formatting of Mathematical Components

This is an example of an equation:

$$
a = 1.
$$

If you want numbered equations use Latex and wrap in the equation environment:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please punctuate 
equations as regular text.

```{comment}
If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
```

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

Theorem-type environments (including propositions, lemmas, corollaries etc.) 
can be formatted as follows:

Example of a theorem:

::: {.Theorem latex=true}
Example text of a theorem
:::

The text continues here. Proofs must be formatted as follows:

Example of a proof:

::: {.proof latex="[Proof of Theorem1]"}
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is 
clear which theorem is being referred to.
:::

